A New Metric for the Analysis of Swarms using Potential Fields

### 2018-08-21 paper.tex
More Minors

\section{Conclusion}\label{Section:MagnitudeDistanceComparison}
%% The two metrics (distance and \textit{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean. The \textit{cohesion/repulsion} metric also allows the state of the swarm to be identified.
%% 
%% The distance metric is based upon the physical \textit{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \textit{interaction} of the agents.
Distance-based metrics measure a swarm's structure in terms of the position of agents. This is independent of the algorithm that produces the swarm's structure. It provides an analysis at a point in time but does not allow the identification of the future behaviour of a swarm as `expanding', `stable' or `contracting'.

@@

The \textit{cohesion/repulsion} metric is based upon the algorithms that have produced the structure and is therefore an analysis of how the algorithms are performing rather than the structure presented. The new metric also provides an indication of a swarm's potential future movement. This is independent of the physical distribution of the agents within the swarm. The metric also provides the same information as the distance metric in terms of agent distribution when the field effects are consistent across all agents i.e. monolithic field effects. This is shown in Fig. \ref{emerge:FIELDFILL-MAG} and \ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same.

The two metrics together provide a deeper evaluation of a swarm's structure and `potential' than either of the metrics alone. Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \textit{cohesion/repulsion metric} falls below zero. This indicates `something' is confining the swarm's distribution as discussed in Section \ref{Section:ApplicationFloodFilling}. This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}

@@

The detection of a swarm's perimeter~\cite{MD:09, MJ:08, ZAPS:07, JG:13}
identifies a subset of a swarm which can be used for coordination. 
Perimeter detection is widely used in swarm coordination~\cite{ZFG:13, AKK:08,
APZDAMC:09, AZDPS:11}. A possible adaptation of this area of research is to
investigate the reduction of the field effects (cohesion and repulsion) of
agents at the perimeter, so increasing agent density in this region. Increasing
the density would allow increased coverage by the swarm at its perimeter. This
could improve the granularity of analysis, e.g. when using sensor arrays at
the frontier of the swarm.
However, this approach leads to swarms of agents having heterogeneous field
effects, as shown in Fig.~\ref{additional:fieldsWork}. Clearly, the
distance-based metric is not useful in the analysis of such swarms. We believe
that the cohesion/repulsion metric introduced in this paper can identify a
swarm's structure even when the swarm is heterogeneous, since the metric is
based directly upon inter-agent field effects rather than the physical
distribution of agents. This will also be the subject of forthcoming work.

>>>>

\section{Conclusion}\label{Section:MagnitudeDistanceComparison}
%% The two metrics (distance and \textit{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean. The \textit{cohesion/repulsion} metric also allows the state of the swarm to be identified.
%% 
%% The distance metric is based upon the physical \textit{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \textit{interaction} of the agents.
Distance-based metrics measure a swarm's structure in terms of the position of agents. This is independent of the algorithm that produces the swarm's structure. It provides an analysis of agent distribution at a point in time but does not allow the identification of the future behaviour of a swarm as `expanding', `stable' or `contracting'.

@@

The new \textit{cohesion/repulsion} metric is based upon the algorithms that have produced the structure and is therefore an analysis of how the algorithms are performing rather than the structure presented. The new metric also provides an indication of a swarm's potential future movement. This is independent of the physical distribution of the agents within the swarm. The metric also provides the same information as the distance metric in terms of agent distribution when the field effects are consistent across all agents i.e. monolithic field effects. This is shown in Fig. \ref{emerge:FIELDFILL-MAG} and \ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same. In addition the new metric provides a mechanism for evaluating swarms that use different algorithms to create a swarming effect and allows the algorithms inter-agent interactions to be compared rather than the resultant structure the algorithm creates.

The two metrics can be used together to provide a deeper evaluation of a swarm's structure and `potential'. Consider the following scenario: the repulsion field is increased but the internal distances do not change; as a result the \textit{cohesion/repulsion metric} falls below zero. This indicates `something' is confining the swarm's distribution as discussed in Section \ref{Section:ApplicationFloodFilling}. This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}

@@

The detection of a swarm's perimeter~\cite{MD:09, MJ:08, ZAPS:07, JG:13}
identifies a subset of a swarm which can be used for coordination. 
Perimeter detection is widely used in swarm coordination~\cite{ZFG:13, AKK:08,
APZDAMC:09, AZDPS:11}. A possible adaptation of this area of research is to
investigate the reduction of the field effects (cohesion and repulsion) on the perimeter
so as to increase the agent density in this region of the swam. This change in the swarm's structure
could improve the granularity of analysis, e.g. when using sensor arrays at
the frontier of the swarm.
However, this approach leads to swarms of agents having heterogeneous field
effects, as shown in Fig.~\ref{additional:fieldsWork}. Clearly, the
distance-based metric is not useful in the analysis of such swarms. We believe
that the cohesion/repulsion metric introduced in this paper can identify a
swarm's structure even when the swarm is heterogeneous, since the metric is
based directly upon inter-agent field effects rather than the physical
distribution of agents. This will also be the subject of forthcoming work.
